Recovery of a deleted deep sequencing data set sheds new light on the early spread of SARS-CoV-2 in Wuhan

### 2021-06-03 paper/paper.tex
first part of recovery of deleted data results

@@

\begin{figure}[]
\centering
\fbox{\includegraphics[trim={0.8in 0.8in 1.1in 1in},clip,width=\linewidth]{figures/pangolin_emails.jpg}}
\caption{Example of the process to delete SRA data.
The image shows an excerpt of e-mails between the lead author of the pangolin coronavirus paper \citet{xiao2020isolation} and SRA staff.
The full e-mails are at \url{https://usrtk.org/wp-content/uploads/2020/12/NCBI-Emails.pdf}.
}
\label{fig:pangolin_emails}
\end{figure}

@@

\subsection{The deleted data set contains sequencing of viral samples collected early in the Wuhan outbreak}
The metadata in the first supplementary table of \citet{farkas2020insights} indicates that the samples in the deleted project PRNJA612766 were collected by Aisu Fu and Renmin Hospital of Wuhan University.
Google searching for these terms revealed the samples were related to a study posted as pre-print on \textit{medRxiv} in early March of 2020~\citep{Wang2020medRxiv}, and subsequently published in the journal \textit{Small} in June of 2020~\citep{Wang2020small}.

The study describes an approach to diagnose infection with SARS-CoV-2 and other respiratory viruses by Oxford Nanopore sequencing.
This approach involved reverse-transcription of total RNA from swab samples, followed by PCR with specific primers that generated amplicons covering portions of the viral genome.
These amplicons were then sequenced on an Oxford Nanopore GridION, and infection with SARS-CoV-2 was diagnosed if the resulting sequencing yielded a sufficiently large number of reads aligning to the SARS-CoV-2 genome.
Importantly, the study notes that this approach yields genetic information about the sequence of the virus as well enabling diagnosis of infection.

@@

\subsection{Recovery of the deleted sequencing data from the Google cloud}

>>>>

\usepackage[htt]{hyphenat}  % https://tex.stackexchange.com/a/543

\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}

@@

\begin{figure}[]
\centering
\fbox{\includegraphics[trim={0.8in 0.8in 1.1in 1in},clip,width=\linewidth]{figures/pangolin_emails.jpg}}
\caption{Example of the process to delete SRA data.
The image shows excerpted e-mails between the lead author of the pangolin coronavirus paper~\citep{xiao2020isolation} and SRA staff.
The full e-mails are at \url{https://usrtk.org/wp-content/uploads/2020/12/NCBI-Emails.pdf}.
}
\label{fig:pangolin_emails}
\end{figure}

@@

\subsection{The deleted data set contains sequencing of viral samples collected early in the Wuhan epidemic}
The metadata in the first supplementary table of \citet{farkas2020insights} indicates that the samples in deleted project PRNJA612766 were collected by Aisu Fu and Renmin Hospital of Wuhan University.
Google searching for these terms revealed the samples were related to a study posted as pre-print on \textit{medRxiv} in early March of 2020~\citep{Wang2020medRxiv}, and subsequently published in the journal \textit{Small} in June of 2020~\citep{Wang2020small}.

The study describes an approach to diagnose infection with SARS-CoV-2 and other respiratory viruses by nanopore sequencing.
This approach involved reverse-transcription of total RNA from swab samples, followed by PCR with specific primers that generated amplicons covering portions of the viral genome.
These amplicons were then sequenced on an Oxford Nanopore GridION, and infection with SARS-CoV-2 was diagnosed if the resulting sequencing yielded a sufficient number of reads aligning to the SARS-CoV-2 genome.
Importantly, the study notes that this approach yields information about the sequence of the virus as well enabling diagnosis of infection.

@@

\subsection{Recovery of deleted sequencing data from the Google cloud} 
As indicated in Figure~\ref{fig:acc_removed}, none of the deleted sequencing runs could be accessed through the SRA's web interface.
In addition, none of the sequencing runs could be accessed using the command-line tools of the SRA toolkit.
For instance, as of June 3, 2020, running the commands \texttt{fastq-dump SRR11313485} or \texttt{vdb-dump SRR11313485} both returned an error message stating ``err: query unauthorized while resolving query within virtual file system module - failed to resolve accession 'SRR11313395'``.

However, the SRA has begun storing all data on the Google and Amazon clouds.
While inspecting the SRA's data access web interface for other sequencing accessions, I noticed that the SRA files were often available from links to the cloud such as \texttt{https://storage.googleapis.com/nih-sequence-read-archive/run/<ACCESSION>/<ACCESSION>} where \texttt{<ACCESSION>} is the accession.

Based on the hypothesis that deletion of sequencing runs by the SRA might not delete files stored on the cloud, I interpolated the cloud URLs for the deleted accessions and tested if they still yielded the SRA files.
This approach was successful; for instance, as of June 3, 2020, going to \url{https://storage.googleapis.com/nih-sequence-read-archive/run/SRR11313485/SRR11313485} downloads the SRA file for accession SRR11313485.
I have digitally archived this URL on the Wayback Machine at \url{https://web.archive.org/web/20210502130820/https://storage.googleapis.com/nih-sequence-read-archive/run/SRR11313485/SRR11313485}.
